### 2013-04-01 trash.tex
Initial commit

\subsection{Software pipelines}
\cite{caporaso2010qiime}
\cite{schloss2009introducing}
\cite{lynch2013axiome}