### 2013-06-05 trash.tex
cranking away...

Substantial attention was given to whether microbial ``types'' exist.
A high-profile paper from Per Bork's group had enterotypes: Bacteroides, Prevotella or Ruminococcus.
\cite{arumugam2011enterotypes}
Appears that there is a real tendency to cluster around Bacteroides or Prevotella \cite{wu2011linking}