### 2013-06-11 trash.tex
-

Substantial attention was given to whether microbial ``types'' exist.
A high-profile paper from Per Bork's group had enterotypes: Bacteroides, Prevotella or Ruminococcus.
\cite{arumugam2011enterotypes}
Appears that there is a real tendency to cluster around Bacteroides or Prevotella \cite{wu2011linking}

\cite{ColeEaRDP2009}
\cite{carma}
\cite{bazinet2012comparative}
A comparative evaluation of sequence classification programs
\cite{brady2011phymmbl}
PhymmBL expanded: confidence scores, custom databases, parallelization and more
\cite{brady2009phymm}
Phymm and PhymmBL: metagenomic phylogenetic classification with interpolated Markov models
\cite{liu2008accurate}
Accurate taxonomy assignments from 16S rRNA sequences produced by highly parallel pyrosequencers
\cite{lan2012using}
Using the RDP Classifier to Predict Taxonomic Novelty and Reduce the Search Space for Finding Novel Organisms
\cite{macdonald2012rapid}
Rapid identification of high-confidence taxonomic assignments for metagenomic data
\cite{maidak1997rdp}
The RDP (ribosomal database project)
\cite{mcdonald2011improved}
An improved Greengenes taxonomy with explicit ranks for ecological and evolutionary analyses of bacteria and archaea
\cite{parks2011classifying}
Classifying short genomic fragments from novel lineages using composition and homology
\cite{phylopythia}
McHardy, A. C.  and Martin, H. G.  and Tsirigos, A.  and Hugenholtz, P.  and Rigoutsos, I.
\cite{munch2008statistical}
Statistical assignment of DNA sequences using Bayesian phylogenetics
\cite{munch2008fast}
Fast phylogenetic DNA barcoding
\cite{rosen2011nbc}
NBC: the Naive Bayes Classification tool webserver for taxonomic classification of metagenomic reads
\cite{segata2011metagenomic}
Metagenomic biomarker discovery and explanation
\cite{segata2012metagenomic}
Metagenomic microbial community profiling using unique clade-specific marker genes

\cite{morlon2011spatial}
Spatial patterns of phylogenetic diversity