### 2014-04-04 reviews/response.txt
Adding decision and start of response.

> Reviews from Prof. Frank (Andy) Anderson
> ----------------------------------------

> Thank you for your Systematic Biology submission. It has been reviewed
> by Associate Editor Prof. Olivier Gascuel and three reviewers. Their
> comments are listed at the end of this letter. The reviewers and the
> AE provide some excellent constructive suggestions that I am sure you
> will appreciate.

> I have little to add that wasn't covered by one or more reviewers or
> Olivier in their fine reviews. As an instructor who has almost given
> up on trying to convince microbiology graduate students that they can
> do better analyses than neighbor joining, I was already sensitive to
> some of these differences, but this review helped me appreciate the
> interface between the two disciplines. I agree with Mooers that the
> level of detail varies across the manuscript, but I mainly noticed
> when a topic was summarized quickly and I didn't really understand it
> (I note that Mooers and I both had a question about how tax2tree
> works).

> I have a few comments and questions that may help clarify some of the
> writing. Altogether, I think the manuscript really requires only
> several minor revisions.

> Like Olivier, I note that some of my comments may overlap those of the
> reviewers.

> 1) Pg. 9: "As described above "metagenome" means that data is sheared
> randomly across the genome rather than amplified from a specific
> location," - I assume you mean DNA is sheared randomly?

> 2) Remember, "data" still is a plural noun, at least under my watch!
> Check for "this data", "data is", "data was", etc

> 3) Pg. 10: I don't really understand what "tax2tree" does, based on
> the description here. Does that matter?

> 4) Pg. 14-15: I also don't really understand what UniFrac is or how it
> works, even after looking at Figure 2 (Olivier made a similar
> comment). Could we get some help here?

> 5) The figures need a bit of work. There seems to be some sort of
> equation at the bottom of Figure 1, but I can't decipher it. What do
> the colors mean in Figure 2? I don't find this to be very clear.

> 6) Pg. 20: "How to formulate a useful notion of independent contrasts
> (Felsenstein 1985) on such an object?" - Awkward wording.

> 7) Pg. 23: I don't understand the distinction between "base
> observations" and sample observations here. Can this be clarified?

> 8) Pg. 24: "Future will continue to bring deeper sequencing on more
> samples" - I guess this should be "The future will continue"?

> My recommendation is acceptance pending minor to major changes.
> Actually, no change asked by the reviewers and me (see below) requires
> major efforts, but we all come from different origins, with different
> points of view and expectations, and thus summing all these demands
> will require a bit of time and attention. Basically, we all ask for
> more details and comments here and there, to make the paper more
> precise, comprehensive and self-contained. I am confident that you'll
> find our suggestions useful. I don't believe that the paper should be
> shortened in some places (or very locally), as suggested by Arne
> Mooers. When submitting your revision, please prepare a cover letter
> containing all the comments by the referees and me, with your
> changes/responses.

> Specific comments, possibly redundant with those from the reviewers,
> using Systematic Biology page numbering:

> P4	these techniques AND could

> P8	(Kuczynski et al. 2010). However, for more subtle information

> P8	"for more subtle information deeper sequencing is required" :
> what do you have in mind here?
> Indeed "distinguish between two rather different samples" is a modest
> goal Actually, it's not clear in this subsection what is meant by
> "Microbial community estimation" (title of the subsection). What is
> the goal here? Identify and classify microbes? Measure their
> abundance? Etc.? A few lines and explanations should suffice.

> P11	What is the level of uncertainty in current bacterial
> phylogenies? For example, one commonly considers that only a few
> unknowns remain in mammals phylogeny; what about bacteria? Are people
> discussing the details or the broad lines? Provide a few illustrative
> examples.

> P13 	"There are many OTU inference methods with various speeds and
> strategies" and the two following paragraphs. Describe in a simple way
> the basic principles of these methods, before discussing their
> relative merits (this remark may apply in other places).

> P14 	You describe PD and suggests that it should be used more often.
> But what are the standard alpha diversity measures, and their
> limitations?

> P15 	Are "partially abundance weighted diversity measures" using some
> phylogenetic approaches? Be a bit more specific when describing and
> discussing methods. Here and in many other places (e.g. UniFrac,
> weighted PD, weighted UniFrac, etc.).

> P15 	"Evans and Matsen (2012) showed that weighted UniFrac" check
> sentence.

> P16 	explanations and Fig. 3 on earth-movers distance are unclear to
> me.

> P17 	"By doing so they have recovered more novel protein families
> than they would have using methods organized by taxonomy". I'm lost
> here, as it is said just above that they have derived a
> "phylogeny-driven genomic encyclopedia of Bacteria and Archaea".

> P23-24 Paragraph "In the same vein This would be a false positive." is
> unclear to me.

> P24 Check "The presently used approach is rarefaction, which means
> uniform sub-sampling, typically without replacement to the number of
> reads in the lowest abundance sample"

> P25 Check "Comparative studies will continue to investigate what
> shapes and is shaped by the microbiome"

> P26 "during the 2013 LIRMM Mathematical and Computational Evolutionary
> Biology workshop", change LIRMM into MCEB

> Fig 2. It took time to me to understand and check on the Web that the
> tree must be rooted, to compute the unifrac distance. This is one
> example, among others, that more details are required when describing
> methods, criteria, measures, etc. Some more figures could help, or one
> appendix with formal definitions of the main concepts and tools. At
> least, details and explanations have to be added in the main text.

> Your reference list is extremely long, mostly due to the fact that you
> are citing papers with a lot of authors. Check Syst Biol
> recommendation to authors (I assume that our editorial office will
> answer and help).

> Reviewer 1, Alexis Stamatakis
> -----------------------------

> Comments:
> This is a well written and in depth review on the usage of
> phylogenetics in human microbiome research, which I enjoyed reading.

> I only have a few minor remarks:

> It might be a consideration to cite our recent paper:

> J. Zhang, P. Kapli, P. Pavlidis, A. Stamatakis: "A General Species
> Delimitation Method with Applications to Phylogenetic Placements". In
> Bioinformatics (2013), 29 (22): 2869-2876.

> It would fit will into section "The role of OTUs".

> Apart from that, I believe that a more critical and slightly extended
> discussion of taxonomic classifiers in the very last part of the paper
> should be included. From my point of view, this is still a mostly
> unresolved problem and the currently best classifier, which you may
> also want to cite,  still shows pretty horrible accuracy
> (http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0049334)
> at the genus and family level.

> I would also appreciate if you could explicitly mention the Gartner
> hype curve in the discussion part and provide an estimate as to where
> we are on this curve at present.

> Some typos:

> page 5 lines 7-8 "these techniques and could be used" there seems to
> be a verb missing.

> page 14 lines 12 Here you may add our recent work on PTP, since it
> strives to conduct OTU clustering and phylogenetic inference in one
> single step/pipeline.

> page 15 l 50: in "distance, is special" seems that it should read
> "distance, it is special"

> page 20: lines 51-56: In fact we are currently developing a tool for
> automatic tree extension as described here. While not published yet,
> it is already available as open-source code at:
> http://sco.h-its.org/exelixis/web/software/put/index.html

> Reviewer 2, Arne Mooers
> -----------------------

> This was an interesting manuscript to read, and I compliment whoever
> thought to review Human Microbiome Research (HMR) for the SB
> readership. Overall, there is a lot of very interesting work reported
> on here, though I found the level of detail to be a bit uneven:
> sometimes we are told perhaps too much, and often, what feels like too
> little, in order to make the connections between HMR and SB. There is
> also seems to be some elision between HMR research and metagenomics
> more generally. That isn't a bad thing, but one does sometimes forget
> that we are focussed on HMR.

> Numbers refer to the SB manuscript pagination, rather than to the
> original latex.

> 4-45 additional to
> 5-8 delete and
> 7-23 delete though

> The section "The Human Microbiome" is quite detailed (discussing mouse
> to human transplants wrt obesity, etc.). Given the intense media
> reporting of the work coming from this area, SB readers may not need
> to be primed of the importance of this field. If this were shortened,
> there could be room to expand some of the technical bits that SB
> readers will want that follow. The last part of the section  - about
> sticking to eubacteria and a non-molecular biology approach do seem
> germane.

> 8-12. these data (I guess we will eventually decide that data is
> singular - up to the editors)

> 8-37++ Whereas Woese ++ This is a bit wordy and unnecessary - we know
> that Woese didn't have PCR

> 10-30 Which "databases" ? - this has been written so far as if it were
> being applied to samples, not databases. I am not clear what the unit
> of observation is here.

> 10-32 We quickly switch here from easy background (nonPCR random
> sequencing) to something brand new (at least to me). "At least some
> portion of the rest of the data can be taxonomically classified
> (methods reviewed by Mande et al. 2012); Treangen et al. (2013) report
> speedups and much higher accuracy when reads are assembled before they
> are classified."
> "Rest of data" refers to what - reads found in a database, or reads
> from a sample? "Accuracy" refers to what? "Speed-ups" refer to what?

> 11-53. Here is the first real point of contact between phylogenetics
> and HMR. But what is an "F-measure of precision and recall" - there is
> no citation.

> 12-28 and 12-32. What is the measure of performance that this refers
> to? How do we decide if a method performs well?

> 12-29. Do readers of SB now what "k-mer composition" is? Is it k-means
> clustering?

> 12-37++. Is the MEGAN approach compositional or phylogenetic? I
> believe it is meant to be a hybrid approach (BLAST being
> compositional, but the taxonomic labeling considered phylogenetic),
> but this could be clarified.

> 13-29. It is not clear what heuristics are required for a cut-off
> method. The author says that he is going to describe the OTU inference
> industry, but the paragraph is really just a list of acronyms, and
> then refers obliquely to a "contingency table" of OTU observations.
> So, for example, what is a notion of goodness of clustering that the
> heuristics are trying to achieve? What are the sorts of species
> divisions found in taxonomies? This is an area where more would be
> helpful - and where a shorter introduction might offer space.

> 14-40. Whereas just about every 16S survey investigation involves an
> OTU-based alpha diversity estimate, only a few involve PD: unweighted
> PD has been applied to some 16S survey data (Lozupone and Knight 2007;
> Costello et al. 2009) and to metagenomic reads in a set of marker
> genes (Kembel et al. 2011) - This is an odd pair of sentences: is the
> L&K the only study that used unweighted PD? If so, in what context?

> 14-46. Check this sentence - perhaps start with "While are commonly
> used, abundance weighted PD measures are not.

> 15-30. Perhaps "In contrast to measures of alpha diversity, measures
> of beta diversity do make use of phylogenetics"  This is actually an
> interesting contrast. Is this just historical contingency, or is there
> a logical reason why the two areas disagree on the use of phylogenetic
> information?

> 15-xx. This seems a strong point of contact between metazoan
> phylogenetics and the microbiome folks. UniFrac is clearly described
> on its own terms, except perhaps where I would have written "with the
> fraction of edges connecting samples that are common to two samples"
> rather than "edges that connect between samples" (which doesn't
> strictly make sense). I believe there is the same measure of
> beta-diversity in the community-phylogenetics literature? What is the
> closest analogue?

> 15-4955. The description of Evans and Matsen seems a bit detailed in
> the context -

> 16-16. This requires another sentence of explanation. From this
> reading, the Unifrac method offers a single distance between a pair of
> samples (a ratio of two numbers of edges), rather than a matrix.

> 16-42..51. This is a short paragraph that would seem to explicitly
> link HMR and SB. Why is it so telegraphic?

> 17-45. This seems an odd heading for this section, which seems to be
> about horizontal gene transfer. Given the amount of technical detail
> elsewhere, why are we not told how horizontal gene transfer is
> inferred in these HMR studies?

> Phylogenetic inference. I found this survey very useful. Is there a
> reason this is not an earlier section of the paper, given these are
> the trees upon which many of the comparisons referred to in previous
> sections are based?

> 19-24. "which has ported in the" Is this the correct structure?

> 20-50. "different parameter regimes" - this phrase is used several
> times in this review. What does it mean here? When would we prefer to
> add sequences to a fixed tree rather than re-infer the totality? Is
> there an analogous situation in SB - i.e. is there something SB can
> learn from HMR here?

> I found the challenges section to be rich and well-written, with
> questions that will resonate with the SB community - e.g. the
> discussion of the proxy problem (how edge lengths predict feature
> diversity). The fact that 16S distance does predict whole genome
> distance is highlighted here - it is presented in previous sections,
> but perhaps it could be given more prominence there, given its
> prominence here. This is a general problem.

> 22-10. "Microbial ecology" Perhaps "and explicitly for human microbial
> ecology" to remind us that HMR is the focus here?

> 23-23. "Whole organism approach"  Will this approach be known to SB
> readers? I cannot imagine what is being done here.

> 25-31. "At least for the next several years we can expect more related
> work." What does this mean? Related to what has been reported above?

> 25-32. The future

> I found the final sentence to leave us dangling. After having reviewed
> this manuscript, I could not say exactly where the intersection of SB
> and HMR will be most fruitful for HMR, nor what in HMR might be
> fruitful for SB. The conclusion seems to suggest that phylogenetics
> would not be used in diagnostics, for example. Will it be important in
> primary research then? In constructing the taxonomies the diagnostics
> may require? Finally, the abstract says the review ends by describing
> "the current challenges to phylogenetics coming from this type of
> work."  My reading of the relevant sections is that there is more
> discussion of challenges to phylogenetic HMR research (sampling
> biases, robustness of UniFrac to incorrect tree reconstruction) than
> to phylogenetics per se. The discussion of how to update known trees
> with new sequences is an exception.

> Reviewer 3, Frederic Delsuc
> ---------------------------

> This manuscript presents a review of how phylogenetic tree-thinking
> and methodology is currently used in the study of the human
> microbiome. The author not only reviews the different ways by which
> phylogenetics is currently percolated into microbial ecology, but also
> gives his personal views on the future phylogenetic challenges
> associated with microbiome studies, and insists on the crucial role
> that molecular phylogeneticists has to play in this quickly expanding
> field. Overall, I enjoyed reading this particularly timely manuscript,
> which I find well written. I only have minor comments/suggestions for
> potential improvement.

> Specific comments:
> 1. I think that it should be made clear, from the very beginning of
> the introduction, the definition of the microbiome in order to avoid
> confusion. Indeed, in the current literature, "microbiota" (i.e. the
> collection of microbes living in association with a given organism)
> and "microbiome" (i.e. the collection of genomes of those microbes)
> are more and more often used interchangeably, even though they are not
> referring to the same organizational level. In that sense, the
> definition of the "human microbiome" given at the top of page 5 ("The
> human microbiome is the collection of microbial organisms that live
> inside of and on the surface of humans.") does not seem to be
> accurate. Please clarify.

> 2. In the section introducing the human microbiome, I think that a
> paragraph on the origin and the way in which the microbiome is
> transmitted vertically between the mother and infant is lacking. The
> recent review on maternal transmission by Funkhouser and Bordenstein
> 2013 is cited later, but it would be nice to mention it on page 5,
> especially in the context of the some examples on the impact of birth
> method (vaginal birth vs. cesarean delivery) on future health of
> newborns.

> 3. On page 6, lines 26-27, it might be worth citing also the recent
> paper by Findley et al. (doi:10.1038/nature12171) on the topography of
> fungal and bacterial on human skin.

> 4. On page 6, line 56, maybe also specify that "survey" is the
> equivalent of barcoding, a term with which the Systematic Biology's
> community is more familiar.

> 5. On page 14, lines 14-19, maybe mention the widespread use of
> "rarefaction curves" to enable the comparison between samples of
> different sequencing depth or methods (e.g. 454 vs. Illumina).

> 6. On page 17, lines 54-55, please add some recent references on the
> current debate around the rooting of the placental tree. I don't think
> that most readers of Syst. Biol. will be familiar with this
> discussion.

> 7. On page 18, lines 26-27, RAxML is also implemented in QIIME.

> 8. On page 20, I think that this might be the place to discuss the
> potential use of phylogenetic networks in microbiome research (see
> Parks and Beiko 2012 Mol Biol Evol).

> 9. On page 24, when citing uBiome, it might also be fair to refer the
> similar crowd-funded American Gut Project.

> Typos:
> P3, L40: I would rather use "species delimitation" instead of "species definition".
> P4, L9: "These techniques and could be used"
> P6, L32: "we will be primarily be describing"
> P9, L29, L30: Maybe replace "diversity" by "variability"
> P18, L32: "a long history as a means to sequentially build"
> P19, L16: "GreenGenes 13_5 alignment has is almost 80% gaps"